---
title: "Vanilla IG tests"
output: html_notebook
---

Compare different IG configurations, with destruction size $d \in \{2, 4, 8, 12\}$:

```{r echo=F}
# loading data
source(here::here("R/read_runs.R"), verbose = F)

runs <- read_runs("rs-ig-destruction-sizes") %>%
  filter(name != "IG1(d=2)") %>%
  filter(name != "IG1MAB")
bounds <- read_csv(here::here("data", 'taillard_instances.csv'))

# runs$name <- factor(runs$name, levels = c(
#   "IG (d=2)", "IG (d=4)", "IG (d=8)", "IG (d=12)", "IG LinUCB", 
# "IG LSPS", "IG MAB", "IG PM", "IG RND", "IG TS"
# ))

bounds <- bounds %>%
  mutate(
    instance = sprintf("taillard_rand_%d_%d_%02d.dat", no_jobs, no_machines, inst_n),
    objective = "MAKESPAN"
  )

results <- runs %>%
  inner_join(bounds, by = c("objective", "instance")) %>%
  mutate(
    max_runtime = no_jobs * no_machines * 60,
    best_fitness60 = map_dbl(runs, ~.x %>% pull(fitness) %>% last),
    best_fitness10 = map2_dbl(runs, max_runtime,
                              ~.x %>% filter(runtime <= .y / 6)
                                  %>% pull(fitness)
                                  %>% last),
    rpd60 = 100 * (best_fitness60 - upper_bound) / upper_bound,
    rpd10 = 100 * (best_fitness10 - upper_bound) / upper_bound
  )

```


```{r}
# View(results %>% select(instance, name, rpd60))
results %>%
  group_by(objective, budget, name) %>%
  ggplot(aes(x = name)) +
  geom_boxplot(aes(y = rpd60))

# results %>% filter(instance == 'taillard_rand_50_20_09.dat', name == 'IG (d=4)') %>% pull(runs) %>% last
```


```{r}
# View(results %>% select(instance, name, rpd60))
results %>%
  group_by(objective, budget, name) %>%
  summarise(
    mean_rpd = mean(rpd60),
    ci = sd(rpd60) * sqrt(1 / (2 * n())),
    n = n()
  ) %>%
  ggplot(aes(x = name)) +
  geom_errorbar(aes(y = mean_rpd, ymin = mean_rpd - ci, ymax = mean_rpd + ci)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# results %>% filter(instance == 'taillard_rand_50_20_09.dat', name == 'IG (d=4)') %>% pull(runs) %>% last
```


## Reward types


IteratedGreedy:
  sol = Init()
  sol' = sol
  sol'' = sol
  while (!StoppingCriterion()) :
    sol' = sol
    sol' = Destruction(sol')
    sol' = Construction(sol')
    sol'' = LocalSearch(sol')
    if (AcceptanceCriterion(sol'', sol))
      sol = sol''



IteratedGreedy:
  sol_0 = Init()
  sol_0' = sol_0
  sol_0'' = sol_0
  while (!StoppingCriterion()) :
    d = OperatorSelection(Reward())
    sol_t' = Construction(Destruction(sol_t, d))
    sol_t'' = LocalSearch(sol_t')
    sol_t+1 = Accept(sol_t'', sol_t)
    Feedback(sol_t, sol_t', sol_t'', sol_t+1)

Reward0: Fintess(sol_t) -  Fitness(sol_t+1)
Reward1: Fitness(sol_t) -  Fitness(sol_t'')
Reward2: Fitness(sol_t') - Fitness(sol_t+1)
Reward3: Fitness(sol_t') - Fitness(sol_t'')


```{r}
reward_types_runs <- read_runs("compare-reward-types") %>%
  filter(map_lgl(runs, ~nrow(.x) > 0)) %>%
 inner_join(bounds, by = c("objective", "instance")) %>%
 mutate(
   max_runtime = no_jobs * no_machines * 60,
   best_fitness60 = map_dbl(runs, ~.x %>% pull(fitness) %>% last),
   best_fitness10 = map2_dbl(runs, max_runtime,
                             ~.x %>% filter(runtime <= .y / 6)
                                 %>% pull(fitness)
                                 %>% last),
   rpd60 = 100 * (best_fitness60 - upper_bound) / upper_bound,
   rpd10 = 100 * (best_fitness10 - upper_bound) / upper_bound
 )

reward_types_runs %>%
  group_by(objective, budget, name) %>%
  summarise(
    mean_rpd = mean(rpd60),
    ci = sd(rpd60) * sqrt(1 / (2 * n())),
    n = n()
  ) %>%
  ggplot(aes(x = name)) +
  geom_errorbar(aes(y = mean_rpd, ymin = mean_rpd - ci, ymax = mean_rpd + ci)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



```{r}
folder <- here::here("runs", "reward-distributions")
experiments <- read_csv(
  file.path(folder, "experiments.csv"),
  col_types =
    cols(
      problem = col_character(),
      type = col_character(),
      objective = col_character(),
      budget = col_character(),
      stopping_criterion = col_character(),
      instance = col_character(),
      name = col_character(),
      mh = col_character(),
      params = col_character(),
      seed = col_double(),
      output = col_character()
    )
  ) %>%
  mutate(runs = map(output, read_csv, 
                    col_types = cols(
                      runtime = col_double(),
                      ig = col_double(),
                      lg = col_double(),
                      il = col_double(),
                      ll = col_double())))


rw_dists <- experiments %>%
  unnest(runs)

rw_dists <- rw_dists %>%
  mutate(
    r0 = lg - ig,
    r1 = lg - il,
    r2 = lg - ig,
    r3 = lg - il
  )

```



```{r}
folder <- here::here("runs", "rewards")
experiments <- read_csv(
  file.path(folder, "experiments.csv"),
  col_types =
    cols(
      problem = col_character(),
      type = col_character(),
      objective = col_character(),
      budget = col_character(),
      stopping_criterion = col_character(),
      instance = col_character(),
      name = col_character(),
      mh = col_character(),
      params = col_character(),
      seed = col_double(),
      output = col_character()
    )
  ) %>%
  mutate(runs = map(output, read_csv, 
                    col_types = cols(
                      runtime = col_double(),
                      initialFitness = col_double(),
                      finalFitness = col_double())))


rewards <- experiments %>%
  unnest(runs)

rewards <- rewards %>%
  mutate(
    reward = finalFitness - initialFitness
  )

```

```{r}

rewards %>%
  filter(name %in% c("IGn (d=2)", "IGn (d=4)", "IGn (d=8)")) %>%
  ggplot() +
  facet_wrap(~instance, scales = "free") +
  geom_histogram(aes(x = reward, fill = name)) +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.position = "bottom")
```


```{r, results='asis'}
rewards %>%
  filter(name %in% c("IGn (d=2)", "IGn (d=4)", "IGn (d=8)")) %>%
  ggplot() +
  facet_wrap(~instance, scales = "free") +
  geom_histogram(aes(x = reward, fill = name)) +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.position = "bottom")
```


```{r}
rewards %>%
  filter(name %in% c("IG (d=2)", "IG (d=4)", "IG (d=8)")) %>%
  sample_n(10000) %>%
  ggplot() +
    facet_wrap(~instance, scales = "free") +
    geom_point(aes(x = runtime, y = reward, color = name)) +
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(),
          legend.position = "bottom")
```



```{r}
rewards %>%
  mutate(g1 = name %in% c("IGn (d=2)", "IGn (d=4)", "IGn (d=8)", "IG MAB")) %>%
  filter(g1) %>%
  ggplot() +
    facet_wrap(~instance, scales = "free") +
    geom_smooth(aes(x = runtime, y = reward, color = name, linetype = g1), method = "lm") +
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(),
          legend.position = "bottom")
```