---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse)
library(here)
source(here("R/read_runs.R"))
```

```{r load, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
runs <- list(
  rs_ig_destruction_sizes = read_runs("rs-ig-destruction-sizes"),
  rs_reward_types = read_runs('best-perturb-fitness'),
  aos_ig = read_runs('aos-ig'),
  #aos_tunning = read_runs('aos-tunning'),
  aos_tunning_race = read_runs('aos-tunning-final-test'),
  aos_tunning_generarization = read_runs('aos-tunning-generarization'),
  aos_148 = read_runs('aos-148-final-comparison'),
  aos_10k_final = read_runs('ig-lsps-aos-tunning-10000evals-final')
)
```


# Introduction

**Objective:** use Automatic Operator Selection (AOS) to augment metaheuristics for flowshop and compare it to the meta-learning approach.

- The IG destruction size seems to control exploration/exploitation, which is the goal of most AOS strategies;
- Assumption: there are differences between different IG destruction sizes.

# Assumption #1: is there any difference between different values of destruction sizes?

```{r}
ig_compare_d_dt <- runs$rs_ig_destruction_sizes %>%
  filter(type == 'PERM', objective == 'MAKESPAN', name %in% c("IG (d=2)", "IG (d=4)", "IG (d=8)"))
```

Setup: `r n_distinct(ig_compare_d_dt$instance)` instances, `r n_distinct(ig_compare_d_dt$seed)` replications on permutation flowshop with makespan objective.

```{r}
compareRpdsCI <- function(runs_dt) {
  plt_dt <- runs_dt %>%
    group_by(problem, type, objective, instance, name, seed) %>%
    filter(fitness == min(fitness))
  plt_dt <- plt_dt %>%
    group_by(problem, type, objective, instance) %>%
    mutate(best_known = min(fitness))
  plt_dt <- plt_dt %>%
    group_by(problem, type, objective, instance, name) %>%
    mutate(rpd = 100 * (fitness - best_known) / best_known)
  plt_dt <- plt_dt %>%
    group_by(problem, type, objective, budget, stopping_criterion, name) %>%
    summarise(
      mrpd = mean(rpd),
      ci = sd(rpd) * sqrt(1 / (2 * n()))
    )
  plt_dt %>%
    ggplot(aes(x = name)) +
    geom_errorbar(aes(y = mrpd, ymin = mrpd - ci, ymax = mrpd + ci, color = paste(problem, type, objective))) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "bottom") +
    theme_bw()
}

```

```{r}
ig_compare_d_dt %>%
  compareRpdsCI()
```


- There are significand differecentes between destruction sizes choices;
- It corroborates with the literature;
- $d = 2$ is too much exploitation and $d = 8$ is too much exploration (further investigation involving mean local optima distances / solution destruction distances)

# Assumption #2: how does it compare to a random approach?

```{r}
runs$rs_ig_destruction_sizes %>%
  filter(type == 'PERM', objective == 'MAKESPAN', name %in% c("IG (d=2)", "IG (d=4)", "IG (d=8)", "IG RND")) %>%
  compareRpdsCI()
```

- Random is worse then the average
- Random might imply too much exploration ($d = 8$ has a bigger impact)

# Can a iteration-based reward be used to inform the best destruction size?

Possible adaptation rewards:

```
  IteratedGreedy:
  sol_0 = Init()
  sol_0' = sol_0
  sol_0'' = sol_0
  while (!StoppingCriterion()) :
    d = OperatorSelection(Reward())
    sol_t' = Construction(Destruction(sol_t, d))
    sol_t'' = LocalSearch(sol_t')
    sol_t+1 = Accept(sol_t'', sol_t)
    Feedback(sol_t, sol_t', sol_t'', sol_t+1)´
```

```
  GG = (sol_t+1 - sol_t ) / sol_t
  GL = (sol_t'' - sol_t ) / sol_t
  GG = (sol_t+1 - sol_t') / sol_t'
  GG = (sol_t'' - sol_t') / sol_t'
```

To test the rewards, all destruction sizes are applied and the best perturbation (BP) for each reward is chosen:

```{r}
runs$rs_reward_types %>%
  filter(type == 'PERM', objective == 'MAKESPAN') %>%
  compareRpdsCI()
```

Comparing to the best fixed choice:

```{r}
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)"))
best_reward <- runs$rs_reward_types %>%
  filter(name != "BP GG")
bind_rows(best_ig, best_reward) %>%
  filter(type == 'PERM', objective == 'MAKESPAN') %>%
  compareRpdsCI()
```

- LL reward is considerably better

# Is there a pattern of destructions sizes over runtime?

```{r}
folder <- here::here("runs", "best-perturb-destruction-size")
runtime_ds <- read_csv(
  file.path(folder, "experiments.csv"),
  col_types =
    cols(
      problem = col_character(),
      type = col_character(),
      objective = col_character(),
      budget = col_character(),
      stopping_criterion = col_character(),
      instance = col_character(),
      name = col_character(),
      mh = col_character(),
      params = col_character(),
      seed = col_double(),
      output = col_character()
    )
  ) %>%
  mutate(runs = map(output, read_delim,
                    delim = ' ',
                    col_names = c('runtime', 'd'),
                    col_types = cols(
                      runtime = col_double(),
                      d = col_double()))) %>%
  unnest(cols = c(runs))


instance_info <- str_match(runtime_ds$instance, '^(.*?)\\_(.*?)\\_(\\d*?)\\_(\\d*?)\\_(\\d*?)\\.dat')
runtime_ds <- runtime_ds %>%
  mutate(
    dist = instance_info[,2],
    corr = instance_info[,3],
    no_jobs = as.integer(instance_info[,4]),
    no_machines = as.integer(instance_info[,5]),
    inst_n = as.integer(instance_info[,6])
  )
```

```{r}
runtime_ds %>%
  sample_n(10000) %>%
  ggplot() +
  facet_grid(no_machines~no_jobs, scales = 'free') +
  geom_point(aes(x = runtime, y = d, color = d))
```


```{r}
runtime_ds %>%
  ggplot() +
  facet_grid(no_machines~no_jobs, scales = 'free') +
  geom_smooth(aes(x = runtime, y = d, color = d), method = 'lm')
```


# Can the LL reward be used to inform the best destruction size?

```{r}
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)"))
bind_rows(best_ig, runs$aos_ig) %>%
  filter(type == 'PERM', objective == 'MAKESPAN') %>%
  compareRpdsCI()
```

# TODO

- Parameter tunning for AOS strategies
- LinUCB on test instances
- Measure the effect when the number of local search steps is 1 or 0


```{r}
runs$aos_tunning %>%
  filter(str_starts(name, 'TS')) %>%
  compareRpdsCI()
```


```{r}
runs$aos_tunning %>%
  filter(str_starts(name, 'FRRMAB')) %>%
  compareRpdsCI() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r}
runs$aos_tunning %>%
  filter(str_starts(name, 'PM')) %>%
  compareRpdsCI() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
runs$aos_tunning %>%
  filter(str_starts(name, 'LIN')) %>%
  compareRpdsCI() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r}
tunned <- c(
  'TS-reward1',
  'FRRMAB-rew1-WindowSize100-Scale0.1-Decay0.9',
  'PM-rew1-rewardTypeextabs -alpha0.2-pmin0.1',
  'LINUCB-rew3-Alpha0.1'
)
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)", "IG RND"))
best_aos <- runs$aos_tunning %>%
  filter(name %in% tunned)
bind_rows(best_ig, best_aos) %>%
  filter(type == 'PERM', objective == 'MAKESPAN') %>%
  compareRpdsCI() +
  coord_flip()

```

```{r}
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)")) ##, "IG RND"))
best_aos <- runs$aos_tunning_race 
bind_rows(best_ig, best_aos) %>%
  filter(type == 'PERM', objective == 'MAKESPAN', no_jobs >= 50, no_machines >= 10) %>%
  compareRpdsCI() +
  coord_flip()
```


```{r}
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)", "IG RND"))
bind_rows(best_ig, runs$aos_148) %>%
  filter(type == 'PERM', objective == 'MAKESPAN', no_jobs >= 50, no_machines >= 10) %>%
  compareRpdsCI() +
  coord_flip()
```

```{r}
runs$aos_tunning_generarization %>%
  compareRpdsCI() +
  coord_flip()
```

```{r}
best_ig <- runs$rs_ig_destruction_sizes %>%
  filter(name %in% c("IG (d=4)", "IG RND"))
bind_rows(best_ig, runs$aos_10k_final) %>%
  filter(type == 'PERM', objective == 'MAKESPAN', no_jobs >= 50, no_machines >= 10) %>%
  compareRpdsCI() +
  coord_flip()
```