---
title: "IG with adaptive operator selection for flowshop problems"
output:
  pdf_document: default
header-includes:
  \usepackage{longtable}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 2, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300, 
                      out.width = "100%",
                      # dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")
library(FlowshopSolveR)
library(tidyverse)
library(here)
library(PMCMRplus)
library(wrapr)

wrangle_perf <- function(dt) {
  dt %>%
    mutate(prob_data = map(problem, ~as_tibble(.x@data))) %>% 
    select(-problem) %>% 
    unnest(prob_data) %>%
    mutate(fitness = map_dbl(result, ~.x$cost))
}


filter_ig_default <- function(config) {
  remove_param <- c(
    "IG.AdaptivePosition.AOS.WarmUp.Strategy",
    "IG.AdaptivePosition.AOS.Options",
    "IG.AdaptivePosition.AOS.Strategy"
  )
  config[(str_detect(names(config), "AOS") | str_detect(names(config), "AdaptiveBestInsertion")) & !(names(config) %in% remove_param)]
}

format_config <- function(config) {
  param_names <- names(config) %>%
    str_remove("^IG\\.AdaptivePosition\\.AOS\\.") %>%
    str_remove("^IG\\.AdaptiveBestInsertion\\.AOS\\.") %>%
    str_remove("^IG\\.AdaptiveBestInsertion\\.") %>%
    str_remove("^(PM|FRRMAB|LINUCB|TS)\\.")
  paste(param_names, config, sep = " = ")
}

load_perf <- function(path) {
  perfs <- read_rds(path)
  perfs %>%
    wrangle_perf()
}

load_perfs <- function(perf_folder) {
  test_problems %>%
    mutate(full_path = file.path(perf_folder, path)) %>%
    select(path, full_path) %>%
    mutate(
      perfs = map(full_path, function(path) {
        tibble(
          algo = dir(path),
          perf = map(file.path(path, algo), load_perf)
        )
      })
    ) %>%
    unnest(perfs) %>%
    unnest(perf) %>%
    group_by(path, id) %>%
    mutate(arfp = 100 * (fitness - min(fitness)) / min(fitness))
}

problems <- read_rds(here("reports", "aos", "data", "train_test_sets_df.rds"))

train_problems <- problems %>%
  filter(set_type == "train")

test_problems <- problems %>%
  filter(set_type == "test")
```

```{r}
exp_folder <- here("reports", "aos", "data", "01-adaptive_destruction_size")
perf_folder <- file.path(exp_folder, "perf")
irace_folder <- file.path(exp_folder, "irace")

algo_perfs <- load_perfs(perf_folder)

algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
algo_perfs %>%
  group_by(algo) %>%
  summarize(
    q00 = quantile(arfp, 0.0),
    q25 = quantile(arfp, 0.25),
    q50 = quantile(arfp, 0.5),
    mean = mean(arfp),
    q75 = quantile(arfp, 0.75),
    q100 = quantile(arfp, 1.00)
  )
```

```{r}
exp_folder <- here("reports", "aos", "data", "01-adaptive_destruction_size")
perf_folder <- file.path(exp_folder, "perf")
irace_folder <- file.path(exp_folder, "irace")

algo_perfs <- load_perfs(perf_folder)

algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
friedmanTest(algo_perfs$arfp, algo_perfs$algo, paste(algo_perfs$id, "-", algo_perfs$seed))
```


```{r}
frdAllPairsNemenyiTest(algo_perfs$arfp, algo_perfs$algo, paste(algo_perfs$id, "-", algo_perfs$seed))
```

```{r}
plot(frdAllPairsNemenyiTest(algo_perfs$arfp, algo_perfs$algo, paste(algo_perfs$id, "-", algo_perfs$seed)))
```


```{r}
algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~objective+type, scale = "free") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```


```{r load}


ig_pos_test <- read_rds(here("reports", "aos", "data", "exp-01-position-adapt", "irace", "ig-pos_test-inf-steps.rds"))

ig_pos_perfs <- ig_pos_test %>% 
  unnest(perf) %>%
  wrangle_perf()

ig_default_perfs <- read_rds(here("reports", "aos", "data", "exp-01-position-adapt", "perf", paste0("ig-def_test.rds"))) %>%
  wrangle_perf()
```


## Proposal

Adapt the IG destruction position during the optimization process in order to focus on more interesting parts of the jobs schedule.

The base IG algorithms uses NEH initialization and (fast) best insertion local search. The adaptation algorithms are:

- Probability matching
- FRRMAB
- LinUCB
- Thompson Sampling

Adicional parameters: reward type and warm-up period.

## Methodology

Tune the adapt parameters and compare with the base algorithm.

## Results

### Experiment 1: Adaptation of destruction postition

Original IG framework:
- Initialization: NEH
- Local search: best insertion
- Perturb: destruction size 4, with adaptive destruction position

Problems:
- 50 and 200 jobs
- random-uniform processing times
- Objective: makespan
- Stopping criterion: $J^2 \times M \times 0.002$ ms  (medium)

Best configurations found by irace (5000 configurations evaluation per adapt strategy):

```{r best-configurations}

ig_pos_test %>%
  select(adapt_strategies, irace_result) %>%
  mutate(
    irace_result = map(irace_result, filter_ig_default),
    irace_result = map(irace_result, format_config)
  ) %>%
  unnest(irace_result) %>%
  knitr::kable("latex")
```

Relative fitness performance:


```{r table-arfp}
ig_pos_perfs %>%
  bind_rows(
    ig_default_perfs %>%
      mutate(adapt_strategies = 'default')
  ) %>%
  group_by(instance) %>%
  mutate(arfp = 100 * (fitness - min(fitness)) / min(fitness)) %>%
  group_by(adapt_strategies) %>%
  summarise(arfp = mean(arfp)) %>%
  knitr::kable("latex")
```

```{r boxplot-compare-ranks}
ig_pos_perfs %>%
  bind_rows(
    ig_default_perfs %>%
      mutate(adapt_strategies = 'default')
  ) %>%
  group_by(instance) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness)) %>%
  ggplot(aes(
    x = adapt_strategies,
    y = arfp
  )) +
  geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot(alpha = .5) +
  scale_y_continuous(trans = "pseudo_log") + 
  coord_trans(y = "pseudo_log")
```

Relative fitness performance per number of jobs:

Statistical test:

```{r friedman}
test_dt <- ig_pos_perfs %>%
  bind_rows(
    ig_default_perfs %>%
      mutate(adapt_strategies = 'default')
  )



friedmanTest(
  test_dt$fitness,
  test_dt$adapt_strategies,
  paste(test_dt$instance, test_dt$seed)
)
```


```{r friendman-all-pairs}
frdAllPairsNemenyiTest(
  test_dt$fitness,
  test_dt$adapt_strategies,
  paste(test_dt$instance, test_dt$seed)
)
```

Conclusions:

- Position adaptation its worse than random positions
- Thompson sampling is the best adaptation strategy
- There are few iterations

### Experiment #2: single step IG-LSPS

Motivation:
- Single step local search leads to more iterations
- LSPS destruction position can be more important
- Extra option allowing destruction from any position

Single-step LSPS framework:
- Initialization: NEH
- Local search: single step of best insertion
- Perturb: destruction size 2 with adaptive position, followed by best insertion local search (also single step)

Problems:
- 100, 200, 300 jobs x 20, 40, 60 machines
- random-uniform processing times
- Objective: makespan
- Stopping criterion: $J^2 \times M \times 0.02$ ms (high)

Best configurations:

```{r}
crossing(
    adapt_strategies = qc(probability_matching,frrmab,linucb,thompson_sampling),
    adapt_type = qc(pos)
  ) %>%
  mutate(
    irace_result_file = here("reports", "aos", "data", 'exp-02-single-step', "irace", 
                             paste0("ig_ss_lsps-", adapt_type, "_", adapt_strategies, ".rds"))
  ) %>%
  filter(file.exists(irace_result_file)) %>%
  mutate(
    irace_result = map(irace_result_file, read_rds),
    irace_result = map(irace_result, removeConfigurationsMetaData),
    irace_result = map(irace_result, ~df_to_character(.x[1,]))
  ) %>%
  select(adapt_strategies, irace_result) %>%
  mutate(
    irace_result = map(irace_result, filter_ig_default),
    irace_result = map(irace_result, format_config)
  ) %>%
  unnest(irace_result)
```


```{r}
test_folder <- here("reports", "aos", "data", 'exp-02-single-step', "perf")

dt <- crossing(
    adapt_strategies = qc(probability_matching,frrmab,linucb,thompson_sampling),
    adapt_type = qc(pos)
  ) %>%
  mutate(
    irace_test_file = file.path(test_folder, adapt_type, adapt_strategies, "test.rds")
  ) %>%
  filter(file.exists(irace_test_file)) %>%
  mutate(perf = map(irace_test_file, read_rds)) %>%
  bind_rows(
    tibble(
      adapt_strategies = "default-ss",
      adapt_type = qc(pos),
      perf = list(read_rds(file.path(test_folder, "default", "test.rds")))
    ),
    tibble(
      adapt_strategies = "default-orig",
      adapt_type = qc(pos),
      perf = list(read_rds(file.path(test_folder, "default-original", "test.rds")))
    )
  ) %>%
  unnest(perf) %>%
  wrangle_perf() %>%
  group_by(instance) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness))

dt %>%
  ggplot(aes(
    x = adapt_strategies,
    y = arfp
  )) +
  geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot(alpha = .5) +
  scale_y_continuous(trans = "pseudo_log") + 
  coord_trans(y = "pseudo_log")
```

```{r friedman-ss}
friedmanTest(
  dt$fitness,
  dt$adapt_strategies,
  paste(dt$instance, dt$seed)
)
```


```{r friendman-all-pairs-ss}
frdAllPairsNemenyiTest(
  dt$fitness,
  dt$adapt_strategies,
  paste(dt$instance, dt$seed)
)
```

### Adaptive best-insertion local search

```{r adapt-ps-boxplots, fig.height=5}
test_folder <- here("reports", "aos", "data", 'exp-03-adaptive_best_insertion', "perf")

dt <- tibble(
    algo = c(# "ig-adapt",
             "ig-def",
             "ig-adapt-replace", 
             "ig-adapt-wo-replace"),
    perf_file = file.path(test_folder, paste0(algo, "-perf.rds"))
  ) %>%
  filter(file.exists(perf_file)) %>%
  mutate(perf = map(perf_file, read_rds)) %>%
  unnest(perf) %>%
  wrangle_perf() %>%
  group_by(instance, objective) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness))

dt %>%
  ggplot(aes(
    x = objective,
    y = arfp,
    fill = algo
  )) +
  # geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot(alpha = .5) +
  facet_wrap(~no_jobs)#+
  # scale_y_continuous(trans = "pseudo_log") #+ 
  # coord_trans(y = "pseudo_log")
```

With parameter tunning:

```{r}
irace_folder <- here("reports", "aos", "data", 'exp-03-adaptive_best_insertion', "irace")

crossing(
  ig_variant = c("ig", "ig-lsps"),
  algo = c('ts', 'pm', 'frrmab', 'linucb')
) %>%
mutate(
  irace_result_file = file.path(irace_folder, 
                        paste0(ig_variant, '-adaptive-bi-', algo, '.rds'))
) %>% 
  filter(file.exists(irace_result_file)) %>%
  mutate(
    irace_result = map(irace_result_file, read_rds),
    irace_result = map(irace_result, removeConfigurationsMetaData),
    irace_result = map(irace_result, ~df_to_character(.x[1,]))
  ) %>%
  select(ig_variant, algo, irace_result) %>%
  mutate(
    irace_result = map(irace_result, filter_ig_default),
    irace_result = map(irace_result, format_config)
  ) %>%
  unnest(irace_result) %>%
  knitr::kable("latex", longtable = TRUE)
  
```


```{r exp-03-adaptive_best_insertion, fig.height=5}
test_folder <- here("reports", "aos", "data", 'exp-03-adaptive_best_insertion', "perf")

dt <- crossing(
    ig_variant = c("ig", "ig-lsps"),
    algo = c('ts', 'pm', 'frrmab', 'linucb')
  ) %>%
  mutate(
    perf_file = file.path(test_folder, 
                          paste0(ig_variant, '-adaptive-bi-', algo, '-tuned-perf.rds'))
  ) %>%
  bind_rows(
    tibble(
     ig_variant = c("ig", "ig-lsps"),
     algo = "def",
     perf_file = file.path(test_folder, 
                           paste0(ig_variant, '-def-perf.rds'))
    )
  ) %>%
  filter(file.exists(perf_file)) %>%
  mutate(perf = map(perf_file, read_rds)) %>%
  unnest(perf) %>%
  wrangle_perf() %>%
  filter(objective == "FLOWTIME") %>%
  group_by(instance, objective) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness))

dt %>%
  ggplot(aes(
    x = objective,
    y = arfp,
    fill = paste(algo)
  )) +
  # geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot() +
  facet_grid(~no_jobs)#+
  # scale_y_continuous(trans = "pseudo_log") #+ 
  # coord_trans(y = "pseudo_log")
```

Statistical comparison:

```{r friedman-exp-03-adaptive_best_insertion}
testdt <- dt %>%
  filter(ig_variant == 'ig-lsps')

friedmanTest(
  testdt$fitness,
  testdt$algo,
  paste(testdt$instance, testdt$seed)
)
```


```{r friendman-all-pairs-exp-03-adaptive_best_insertion}
frdAllPairsNemenyiTest(
  testdt$fitness,
  testdt$algo,
  paste(testdt$instance, testdt$seed)
)
```


### Adaptive destruction size:

- Single step
- Acceptance criterion: better

```{r exp-04-adaptive_destruction_size, fig.height=5}
test_folder <- here("reports", "aos", "data", 'exp-04-adaptive_destruction_size', "perf")

dt <- crossing(
    ig_variant = c("ig", "ig-lsps"),
    algo = c('ts', 'pm', 'frrmab', 'linucb')
  ) %>%
  mutate(
    perf_file = file.path(test_folder, 
                          paste0(ig_variant, '-adaptive-ds-', algo, '-tuned-perf.rds'))
  ) %>%
  bind_rows(
    tibble(
     ig_variant = c("ig", "ig-lsps"),
     algo = "def",
     perf_file = file.path(test_folder, 
                           paste0(ig_variant, '-def-perf.rds'))
    )
  ) %>%
  filter(file.exists(perf_file)) %>%
  mutate(perf = map(perf_file, read_rds)) %>%
  unnest(perf) %>%
  wrangle_perf() %>%
  filter(objective == "FLOWTIME") %>%
  group_by(instance, objective) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness))

dt %>%
  ggplot(aes(
    x = objective,
    y = arfp,
    fill = paste(algo)
  )) +
  # geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot() +
  facet_grid(~ig_variant)#+
  # scale_y_continuous(trans = "pseudo_log") #+ 
  # coord_trans(y = "pseudo_log")
```



### Adaptive destruction size:

- Single step
- Acceptance criterion: better

```{r exp-05-adaptive_size_position, fig.height=5}
test_folder <- here("reports", "aos", "data", 'exp-05-adaptive_size_position', "perf")

dt <- crossing(
    ig_variant = c("ig", "ig-lsps"),
    algo = c('ts') #, 'pm', 'frrmab', 'linucb')
  ) %>%
  mutate(
    perf_file = file.path(test_folder, 
                          paste0(ig_variant, '-adaptive-ds-', algo, '-tuned-perf.rds'))
  ) %>%
  bind_rows(
    tibble(
     ig_variant = c("ig", "ig-lsps"),
     algo = "def",
     perf_file = file.path(test_folder, 
                           paste0(ig_variant, '-def-perf.rds'))
    )
  ) %>%
  filter(file.exists(perf_file)) %>%
  mutate(perf = map(perf_file, read_rds)) %>%
  unnest(perf) %>%
  wrangle_perf() %>%
  filter(objective == "FLOWTIME") %>%
  group_by(instance, objective) %>%
  mutate(arfp = 100 * ( fitness - min(fitness)) / min(fitness))

dt %>%
  ggplot(aes(
    x = objective,
    y = arfp,
    fill = paste(algo)
  )) +
  # geom_violin(draw_quantiles = T, alpha = .5) +
  geom_boxplot() +
  facet_grid(~ig_variant)#+
  # scale_y_continuous(trans = "pseudo_log") #+ 
  # coord_trans(y = "pseudo_log")
```


## Testing on different problems


```{r}
exp_folder <- here("reports", "aos", "data", "01-adaptive_best_insertion")

algo_perfs <- load_perfs(
  file.path(exp_folder, "perf"), 
  problems %>% filter(set_type == "test")
)


algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_boxplot() +
  facet_wrap(~path, scales = "free", nrow = 3) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}

train_folder <- file.path(exp_folder, "irace")

load_config <- function(path) {
  read
}

problems %>%
  filter(set_type == "train") %>%
  mutate(full_path = file.path(train_folder, path)) %>%
  select(path, full_path) %>%
  mutate(algo = map(full_path, dir)) %>%
  unnest(algo) %>%
  filter(!str_detect(algo, "_log|_rec")) %>%
  mutate(
    config = map(
      file.path(full_path, algo), 
      compose(
        read_rds,
        ~.x[1,],
        removeConfigurationsMetaData,
        df_to_character,
        filter_ig_default,
        format_config,
        .dir = "forward"
      )
    )
  ) %>%
  unnest_longer(config) %>%
  select(path, algo, config) %>%
  knitr::kable("latex", longtable = TRUE)
```



```{r fig.asp=2}
algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_boxplot() +
  facet_wrap(~path, scales = "free", nrow = 3) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```



```{r}
algo_perfs %>%
  ggplot(aes(x = algo, y = arfp)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r friedman-exp-03-adaptive_best_insertion}
testdt <- algo_perfs %>%
  group_by(id, path, algo) %>%
  summarise(arfp = mean(arfp), .groups = "drop")

friedmanTest(
  testdt$arfp,
  testdt$algo,
  paste(testdt$id, testdt$path)
)
```

```{r friedman-exp-03-adaptive_best_insertion}
testdt <- algo_perfs %>%
  group_by(id, path, algo) %>%
  summarise(arfp = mean(arfp), .groups = "drop")

frdAllPairsExactTest(
  testdt$arfp,
  testdt$algo,
  paste(testdt$id, testdt$path)
)
```
